---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# vecrep

<!-- badges: start -->
<!-- badges: end -->

vecrep provides `rep_altrep()`, an ALTREP alternative to `base::rep()`
for numeric vectors. Rather than duplicating data immediately, it stores a
compact reference to the original vector and only expands it if a write forces
materialisation. This makes it well-suited to vectors with many repetitions,
especially if the reference vector is a regular sequence represented with 
ALTREP. ALTREP sequences can be combined with ALTREP replicates to create 
repeating regular sub-sequences.

## Installation

```r
# install.packages("remotes")
remotes::install_github("mitchelloharawild/vecrep")
```

## Usage

```{r example}
library(vecrep)

x <- as.numeric(1:5)

# Create a repeated vector — no extra allocation
y <- rep_altrep(x, times = 4)

length(y)   # 20
y[1:10]     # reads directly from x
sum(y)      # aggregates stay lazy too
```

Read operations (`[`, `sum()`, `mean()`, `anyNA()`) work directly on the
parent vector without expanding it. The full vector is only materialised on the
first write, and copy-on-write ensures the parent is never modified.

```{r materialise}
parent <- as.numeric(1:5)
y <- rep_altrep(parent, 3)

y[1] <- 999   # triggers expansion

parent        # unchanged
y[1:6]
```

## Caveats

- **Numeric only.** Integer, character, and list inputs are rejected.
- **Serialisation materialises.** `saveRDS()` expands the vector and emits a
  warning; the round-tripped object is correct but no longer compact.
- **Sorting expands the internal buffer**, though the ALTREP shell persists.
