---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# vecrep

<!-- badges: start -->
[![R-CMD-check](https://github.com/mitchelloharawild/vecrep/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/mitchelloharawild/vecrep/actions/workflows/R-CMD-check.yaml)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- [![CRAN status](https://www.r-pkg.org/badges/version/vecrep)](https://CRAN.R-project.org/package=vecrep) -->
<!-- badges: end -->

vecrep provides `rep_altrep()`, an ALTREP alternative to `base::rep()`
that works with most vector types. Rather than duplicating data immediately, 
it stores a compact reference to the original vector and only expands it if a
write forces materialisation. This makes it well-suited to vectors with many
repetitions, especially if the reference vector is a regular sequence
represented with ALTREP. ALTREP sequences can be combined with ALTREP
replicates to create repeating regular sub-sequences.

## Installation

```r
# install.packages("remotes")
remotes::install_github("mitchelloharawild/vecrep")
```

## Usage

```{r example}
library(vecrep)

x <- as.numeric(1:5)

# Create a repeated vector — no extra allocation
y <- rep_altrep(x, times = 4)

length(y)   # 20
y[1:10]     # reads directly from x
sum(y)      # aggregates stay lazy too
```

Read operations (`[`, `sum()`, `mean()`, `anyNA()`) work directly on the
parent vector without expanding it. The full vector is only materialised on the
first write, and copy-on-write ensures the parent is never modified.

```{r materialise}
parent <- as.numeric(1:5)
y <- rep_altrep(parent, 3)

y[1] <- 999   # triggers expansion

parent        # unchanged
y[1:6]
```

The `each` argument repeats each element in turn before moving to the next,
matching the behaviour of `base::rep(..., each = n)`:

```{r each}
x <- as.numeric(1:3)

# Each element repeated 3 times: 1 1 1 2 2 2 3 3 3
rep_altrep(x, each = 3)

# times and each can be combined — each is applied first, then times repeats the result
rep_altrep(x, times = 2, each = 3)
```

## Supported types

`rep_altrep()` supports most vector types:

```{r types}
# integer
rep_altrep(1L:3L, 3L)

# logical
rep_altrep(c(TRUE, FALSE, NA), 2L)

# complex
rep_altrep(c(1+1i, 2+2i), 4L)

# raw
rep_altrep(as.raw(c(0x01, 0x02, 0x03)), 2L)

# character
rep_altrep(c("foo", "bar", "baz"), 3L)

# list
rep_altrep(list(1L, "a", TRUE), 2L)
```

Classed vectors such as `factor`, `Date`, and `POSIXct` are handled
transparently. The class and relevant attributes (e.g. `levels` for factors)
are preserved on the ALTREP object without forcing materialisation, so S3
dispatch works as expected:

```{r classed}
# factor: levels preserved without expansion
f <- rep_altrep(factor(c("cat", "dog", "cat")), 3L)
class(f)
levels(f)
table(f)

# Date
d <- rep_altrep(as.Date("2024-01-01") + 0:2, 2L)
class(d)
d

# POSIXct
p <- rep_altrep(as.POSIXct("2024-01-01") + 0:2, 2L)
class(p)
```

Named vectors keep their names lazily — the names are themselves stored as an
ALTREP character vector rather than being eagerly expanded:

```{r names}
x <- c(a = 1.0, b = 2.0, c = 3.0)
y <- rep_altrep(x, 3L)
names(y)
```

## Caveats

- **Serialisation materialises.** `saveRDS()` expands the vector; the
  round-tripped object is correct but no longer compact.
- **Sorting expands the internal buffer**, though the ALTREP shell persists.

## Acknowledgements

The initial codebase was adapted from Gabriel Becker's
[vectorwindow](https://github.com/ALTREP-examples/vectorwindow/) example,
presented in his Bioconductor Developers Forum talk
([YouTube](https://www.youtube.com/watch?v=8i7ziLqsE2s)).

Substantial proportions of this package were developed in tandem with
[Claude Sonnet 4.6](https://www.anthropic.com/claude) (Anthropic). All code has been reviewed and guided by humans.
